\documentclass[11pt]{amsart}
\usepackage{amssymb,amsthm,amsmath,amstext}
\usepackage{mathdots}

\usepackage[left=1.2in,right=1.2in,top=1.5in,bottom=1.5in,nohead,nofoot]{geometry}

\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\F}{\mathbb{F}}

\pagestyle{empty}

\newcounter{problem}
\setcounter{problem}{0}

\newenvironment{problem}[1][]%
{%
\stepcounter{problem} \vspace{.2cm} \noindent {\bf \arabic{problem}.} {\it #1}~ %
}{%
\vspace{.2cm}%
} 

\renewcommand{\theenumi}{\alph{enumi}}

\begin{document}

\thispagestyle{empty}

\title{} 

%\date{\today}

\noindent\sc
Dartmouth College Department of Mathematics \\ \bf
Math 75 Cryptography \\ \rm
Spring 2022\\[2ex]
Problem Set \# 4 (upload to Canvas by Friday, April 29, 10:10 am EDT)

\bigskip

\noindent {\bf Problems:}

\begin{problem} 
Let 
\[
A=\begin{pmatrix}
1 & 2 & 3 \\
4 & 5 & 6 \\
7 & 8 & 10 
\end{pmatrix}. \]
For which $n$ is the matrix $A$ invertible over $\Z/n\Z$?  Find its inverse if $n=100$.
\end{problem}

\begin{problem}
Alice uses the Hill cipher, encrypting the plaintext
\begin{center}
\textsf{Consistency is the last refuge of the unimaginative}
\end{center}
to get the ciphertext
\begin{center}
\texttt{voqimugocogmttfkxvlvdynhawugtfrsksoizgaanlygk}
\end{center}
to send to Bob using blocks of size $m=3$ (and $n=26$).  Playing the
role of Eve, hack Alice's encryption key $A \in M_3(\Z/26\Z)$.  The
matrix key spells out a keyword: what is it?

After you find the key,
notice that Alice has not followed the protocol correctly.  Explain
why, then find two plaintexts that encrypt to the same ciphertext
using Alice's key.
\end{problem}

\begin{problem}
The Hill cipher succumbs to a known plaintext attack if sufficiently many plaintext-ciphertext pairs are known.
It is even easier to break the cipher if Eve can trick Alice into encrypting a chosen plaintext,
a \emph{chosen plaintext attack}.
Describe such an attack.
\end{problem}

\begin{problem}
Let $n \in \Z_{>0}$.  We consider the row-reduction algorithm over
$\Z/p\Z$ where $p$ is prime.  Recall that every nonzero element in
$\Z/p\Z$ has a multiplicative inverse, so that $\F_p=\Z/p\Z$ is a
\emph{field} and all the methods of linear algebra apply.

Find an explicit polynomial $f(x) \in \Q[x]$ of degree $3$ such that
no more than $f(n)$ operations in $\F_p$ are required by the
row-reduction algorithm for computing the determinant of a matrix in
$M_n(\F_p)$.  How many of these operations are inversions of nonzero scalars?
\end{problem}

\begin{problem}
Decrypt the message
\begin{center}
\texttt{CLV SSH = MMBVC RDMVE PFZII EAVYS XFTHS FNMOB RRPDH VBSQH}
\end{center}
with the following Enigma settings: 
\begin{verbatim}
Walzenlage (Rotors): I V III
Ringstellung (Ring setting): 13 06 24
Steckerverbindungen (Plug connections): AU PB EF IQ RH ZL DT MS CG KN
Kenngruppen: KIJ TFR BVC ZAE
\end{verbatim}
\emph{[Hint: The message is in German!]}
\end{problem}

\begin{problem}
Read Section 15.1 (pages 368--376) of \emph{The Pleasures of Counting}
by Koerner (posted on Canvas) and do Exercises 15.1.1--15.1.3.
\end{problem}

\vfill

\end{document}

