\begin{problem}
Let $n \in \Z_{>0}$.  We consider the row-reduction algorithm over
$\Z/p\Z$ where $p$ is prime.  Recall that every nonzero element in
$\Z/p\Z$ has a multiplicative inverse, so that $\F_p=\Z/p\Z$ is a
\emph{field} and all the methods of linear algebra apply.

Find an explicit polynomial $f(x) \in \Q[x]$ of degree $3$ such that
no more than $f(n)$ operations in $\F_p$ are required by the
row-reduction algorithm for computing the determinant of a matrix in
$M_n(\F_p)$.  How many of these operations are inversions of nonzero scalars?
\end{problem}

\begin{Answer}
Suppose we have an $n$ by $n$ matrix $A$ in $M_n(\F_p)$.

\noindent
We can find the determinant of $A$ by row-reducing $A$ into a triangular matrix 
then finding the product of the pivot-position elements.

\noindent
We first remove the lower non-diagonal elements in the first column,
then the second, and so on up to the $n-1$th column.
The $n$'th column's diagonal element, being in the last row in the matrix,
has no rows below it.

\noindent
This step involves:
\begin{itemize}
  \item Finding the inverse of the diagonal element. This is a total of $n-1$ inversions
    ($1$ per column, over the $n-1$ columns).
  \item Adding a multiple of the row to \textit{all the rows below the current element
    that have a non-zero element in the column of interest},
    such that we cancel out the nonzero value. This involves a multiplication and an addition.
\end{itemize}

\noindent
This is a total of $n - 1$ inversions $+ \sum_{i=1}^{n-1} i$ row additions,
which add up to 

$n - 1 \mathbf{\;inversions}$ 
$+2 \cdot \frac{n(n-1)}{2} = n(n-1) \mathbf{\;row \;multiplications\; and\; additions}$.

\noindent
However, each row multiplication or addition requires doing the operation
on each of the $n$ elements of the row.

\noindent
Thus, the total comes to $n - 1 +
n (n-1)\cdot n = \frac{n^3 - n^2}{2} + n - 1$.

\noindent
Finally, we need to multipy the $n$ diagonal elements to find the determinant,
making a total of 

\zaff{$n^3 - n^2+ 2n - 1$} integer operations.

\noindent
Of these operations, \crim{$n - 1$} are inversions of nonzero scalars, 
and \crim{$n^3 - n^2+ n$} are integer additions and multiplications.

\end{Answer}
