\begin{problem}
Consider the affine cipher with $\calP=\calC=\Z/n\Z$.  
\begin{enumerate}\renewcommand{\itemsep}{3mm}
\item Suppose $n=541$ and we take the key $(a,b)=(34,71)$.  Encrypt the plaintext $m=204$, and decrypt the ciphertext $c=431$.
\begin{Answer}
\noindent
The encryption of $m=204$ is $515$
\begin{align*}
  c &= a \cdot p + b\\
  &= 34 \cdot 204 + 71\\
  &= 7007\\
  &\equiv 515 \crim{\pmod{541}}
\end{align*}

\noindent
The decryption of $c=431$ is $297$
\begin{align*}
  c &\equiv a \cdot p + b \crim{\pmod n}\\
  431 &\equiv 34p + 71 \crim{\pmod{541}}\\
  360 &\equiv 34p \crim{\pmod{541}}\\
  p &= \frac{360 + 541k}{34}  \hspace{15pt}\mid p,k \in \Z^+\\
\end{align*}
\noindent
We need to find a value for  $k$ such that $34 | (541k + 360)$.
This is easily calculated to be $k=18$.
\begin{align*}
  p &= \frac{360 + 541\cdot 18}{34}\\
  p &= 297
\end{align*}
\end{Answer}
\newpage
\item Eve intercepts a ciphertext from Alice
and through espionage she learns that the letter
$x \in \calP$ is encrypted as $y \in \calC$ in this message.
Show that Eve can decrypt the message using $O(n)$ trials.  
\begin{Answer}

Suppose Eve knows that a letter $x \in \calP$ is encrypted as
$y \in \calC$ in the message.

\noindent
Then, Eve knows that \crim{$a \cdot x + b \pmod n \equiv y \pmod n$} for some $a, b \in \Z/n\Z$,
where $(a, b)$ are the keys of the Affine Cipher.

\begin{align*}
  ax + b &\equiv y \pmod n\\
  ax + b &\equiv y\\
  ax &\equiv y - b\\
\end{align*}

\noindent
Eve can safely assume that $0 \le b \le n-1$ 
(since adding any number $x \ge n$ is equivalent to adding
$x \mod n$). Eve can therefore iterate through all the $n$ possible values of $b$ and
test the matching value for $a$. She is guaranteed to find the real key.

\end{Answer}
\item Now suppose that (contrary to Kerckhoffs's principle) the integer $n$ is not public knowledge.
Is the affine cipher still vulnerable if Eve manages to steal a plaintext/ciphertext pair?
How might Eve break the system?

\begin{Answer}
Without knowing $n$, the problem becomes much harder to break.
However, if Eve knows at least $3$ different plaintext/ciphertext pairs,
she can use them to guess a value for $n$.

\noindent
Say, for instance, $p_1, p_2, p_3$ are the plaintexts and $c_1, c_2, c_3$ are the ciphertexts:

\begin{align*}
  c_1 \equiv a \cdot p_1 + b \pmod n &\Rightarrow a \cdot p_1 + b - c_1 \equiv 0 \pmod n\\
  c_2 \equiv a \cdot p_2 + b \pmod n &\Rightarrow a \cdot p_2 + b - c_2 \equiv 0 \pmod n\\
  c_3 \equiv a \cdot p_3 + b \pmod n &\Rightarrow a \cdot p_3 + b - c_3 \equiv 0 \pmod n\\
  \\
  \begin{bmatrix}
    c_1 & p_1 & 1\\
    c_2 & p_2 & 1\\
    c_3 & p_3 & 1
  \end{bmatrix} \cdot \begin{bmatrix}
    1\\
    -a\\
    -b
  \end{bmatrix} &\equiv \begin{bmatrix}
    0\\
    0\\
    0
  \end{bmatrix} \pmod n\\
  \\
  \begin{vmatrix}
    c_1 & p_1 & 1\\
    c_2 & p_2 & 1\\
    c_3 & p_3 & 1
  \end{vmatrix} &\equiv 0 \pmod n\\
  \\
  \color{crimson}
  c_1(p_2-p_3) - c_2(p_1-p_3) + c_3(p_1 - p_2) &\equiv 0 \pmod n \\
  \color{black} \\
\end{align*}

\noindent
Since $\{c_1, c_2, c_3\}$ and the corresponding plaintexts $\{p_1, p_2, p_3\}$ are all known,
Eve can find a value congruent to $0$ in $\Z/n\Z$ and use it to find $n$,
after which she can easily crack the encryption.

\end{Answer}
\end{enumerate}
\end{problem}

